Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2015-10-24 RMS/submission.tex
fixed some typos

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
modeling language and analyzing tool for real-time systems, to model a
realistic implementation of RMS and then verify some desired
properties on the built model. Based on a realistic implementation,
our model extends the standard one proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by allowing computation
requirements of tasks to be non-constant, and by considering overhead
and other details of the enviroments.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
%\subsection{Formulation of the Standard Setting}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated simultaneously. This means that jobs corresponding to
task $\tau_i$ are initiated at times $kT_i$ with integers $k\ge
0$. Deadlines consist of runability constraints only: the job
initiated at time $kT_i$ has $(k+1)T_i$ as its deadline, which is the
initiation time of the next job. The RMS algorithm chooses the
labeling such that $T_1\le T_2\le \ldots \le T_n$. Consequently,
$\tau_i$ receives priority $i$, assuming smaller numbers have higher
priorities. The following assumptions are made:

@@

In this paper, we only consider Real-Time Maude's \emph{linear
  temporal logic model checker}, which analyzes whether \emph{each}
behavior satisfies a temporal logic formula. \emph{State propositions}
are terms of sort \verb|Prop|. Their symantics is defined by equations
of the form $\texttt{ceq } statePattern \texttt{ |= } prop \texttt{ =
} b \texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating
that $prop$ evaluates to $b$ in states which are instances of
$statePattern$ provided the condition $cond$ holds. These equations
together define $prop$ to hold in all states $t$ that make $t \texttt{
  |= } prop$ evaluate to \verb|true|. A temporal logic \emph{formula}
is constructed by state propositions and temporal logic operators such
as \verb|True|, \verb|False|, \verb|~|(negation), \verb|/\|,
\verb|\/|, \verb|->|(implication), \verb|[]|(``always''),
\verb|<>|(``eventually'') and \verb|U|(``until''). Real-Time Maude
supports both \emph{timed} and \emph{untimed LTL model checking}. The
untimed model checking command
\begin{alltt}
  (mc \(t\) |=u \(formula\) .)
\end{alltt}
checks whether the temporal logic $formula$ holds in all behaviors
starting from the initial state $t$, \emph{with no time limit}.

@@

\section{Formal Model of the Implementation}
\label{s:formalism}
We model the implementation under the following assumptions:

@@

(A2) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instanly.

@@

In general, completeness of untimed model checking cannot be achieved
for any systems, any time sampling stategies and any
properties. However, \"Olveczky and Meseguer proved the completeness
of untimed temporal logic model checking, under the maximal time
sampling strategy, for a large class of real-time
systems~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{theorem}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{t:completeness}
Given a \emph{time-robust} real-time rewrite theory $\cal R$, a set
$P$ of \emph{tick-stabilizing} atomic propositions, an LTL formula
$\Phi$ (excluding the \emph{next} operator $\bigcirc$) whose atomic
propositions are contained in $P$. The untimed temporal logic model
checking verifying $\Phi$ is \emph{complete} under the maximal time
sampling strategy.
\end{theorem}

>>>>

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
modeling language and analyzing tool for real-time systems, to model a
realistic implementation of RMS and then verify some desired
properties on the built model. Based on a realistic implementation,
our model extends the standard one proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by allowing computation
requirements of tasks to be non-constant, and by considering overhead
and other details of the environments.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
%\subsection{Formulation of the Standard Setting}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated simultaneously. This means that jobs corresponding to
task $\tau_i$ are initiated at times $kT_i$ with integers $k\ge
0$. Deadlines consist of runnability constraints only: the job
initiated at time $kT_i$ has $(k+1)T_i$ as its deadline, which is the
initiation time of the next job. The RMS algorithm chooses the
labeling such that $T_1\le T_2\le \ldots \le T_n$. Consequently,
$\tau_i$ receives priority $i$, assuming smaller numbers have higher
priorities. The following assumptions are made:

@@

In this paper, we only consider Real-Time Maude's \emph{linear
  temporal logic model checker}, which analyzes whether \emph{each}
behavior satisfies a temporal logic formula. \emph{State propositions}
are terms of sort \verb|Prop|. Their semantics is defined by equations
of the form $\texttt{ceq } statePattern \texttt{ |= } prop \texttt{ =
} b \texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating
that $prop$ evaluates to $b$ in states which are instances of
$statePattern$ provided the condition $cond$ holds. These equations
together define $prop$ to hold in all states $t$ that make $t \texttt{
  |= } prop$ evaluate to \verb|true|. A temporal logic \emph{formula}
is constructed by state propositions and temporal logic operators such
as \verb|True|, \verb|False|, \verb|~|(negation), \verb|/\|,
\verb|\/|, \verb|->|(implication), \verb|[]|(``always''),
\verb|<>|(``eventually'') and \verb|U|(``until''). Real-Time Maude
supports both \emph{timed} and \emph{untimed LTL model checking}. The
untimed model checking command
\begin{alltt}
  (mc \(t\) |=u \(formula\) .)
\end{alltt}
checks whether the temporal logic $formula$ holds in all behaviors
starting from the initial state $t$, \emph{with no time limit}.

@@

\section{Formal Modeling of the Implementation}
\label{s:formalism}
We model the implementation under the following assumptions:

@@

(A2) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly.

@@

In general, completeness of untimed model checking cannot be achieved
for any systems, any time sampling strategies and any
properties. However, \"Olveczky and Meseguer proved the completeness
of untimed temporal logic model checking, under the maximal time
sampling strategy, for a large class of real-time
systems~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{theorem}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{t:completeness}
Given a \emph{time-robust} real-time rewrite theory $\cal R$, a set
$P$ of \emph{tick-stabilizing} atomic propositions, an LTL formula
$\Phi$ (excluding the \emph{next} operator $\bigcirc$) whose atomic
propositions are contained in $P$. The untimed temporal logic model
checking verifying $\Phi$ is \emph{complete} under the maximal time
sampling strategy.
\end{theorem}
